---
title: "ajfhelpR_bools v1"
author: "Andrew Frick"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{str_comb_X Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(tidyverse)
library(ajfhelpR)
knitr::opts_chunk$set(echo = TRUE)
```

One of the first "oops" mistakes I had in R had to do with T | NA == NA instead
of false. While there are times when you don't want to coerce that to false, I
decided to create a new set of boolean operators for this case.

First we'll look at the new `equal to` operator `%==%` by creating a data frame of 
boolean combinations

```{r}
#Create data frame/tibble with combination of all bools

x = c(T,F,NA)

bools = as.tibble(expand.grid(x,x))

equals = bools %>% 
  mutate(`==` = (Var1 == Var2),
         `%==%` = (Var1 %==% Var2))
```

Below you can see the difference in the output for the original `==` and the 
`%==%` extension.

```{r}
equals
```

Here we'll create another tibble containing the 'venn diagram' boolean operators

1. `|` vs `%or%`
2. `&` vs `%&%`

```{r}

Venns = bools %>% 
  mutate(`|` = (Var1 | Var2),
         `%or%` = (Var1 %or% Var2),
         `&` = (Var1 & Var2),
         `%&%` = (Var1 %&% Var2))

```


```{r}
Venns
```

The equality operator also allows comparison of a single value vs a vector, it
can be used for subsetting operations. 

x[(condition)] returns all values where (condition) is not false. NA is 
not interpreted as false by R, so any rows with missing condition will be also
be returned. To ensure your subset only returns those that are true, we can use
the `%==%` operator to coerce those to false.

```{r}
set.seed(12345)

y = 'blue'

colors = c('blue', 'red', 'yellow', NA)

colorsamp = sample(colors, size = 10, prob = c(0.4,0.2,0.1,0.3),
                   replace = T)


y  ==  colorsamp
y %==% colorsamp

colorsamp[colorsamp == y]

colorsamp[colorsamp %==% y]

```

The `%==%` operator is especially useful for `ifelse` statements. Let's say we 
change all y to 0 if x is not equal to 'a'

```{r}
set.seed(12356)

tib = tibble(
  x = sample(c("a", "b", NA), size = 10, replace = T),
  y = sample(1:10, size = 10, replace = T)
)

tib

```
We see there are 3 rows with `NA`, and depending on how we wish to view the data
in context, we could consider these to be 1. "not equal to a" or 2. "unknown." 

```{r}

tib = tib %>% 
  mutate(`y %==%` = ifelse(x %==% 'a', x, 0),
         `y ==`   = ifelse(x  ==  'a', x, 0))

tib
```
The `==` operator perceives `x == NA` as missing and outputs a value of missing
for the new column. The `%==%` opertor coerces those missing booleans to F, and
outputs the value of 0. Either of these approaches can be correct based on the
context of the individual problem. But I've found in data that I work with, 
`%==%` tends to be more useful.
