---
title: "summ_X Functions WIP v1"
author: "Andrew Frick"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summ_X Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(ajfhelpR)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Welcome to my vignette for the `summ_X` functions. Currently there are 3, though I only really use 2 of them. First off, I recommend installing the following packages which I'll be using extensively in all my guides. And let's be honest, they're just pretty dang awesome.

```{r, include = T, echo = T}
## Packages used for demonstration

# install.packages("tidyverse")
# install.packages("kableExtra")
# install.packages("pander")
# install.packages("stringr")
# install.packages("knitr")

library(tidyverse)
library(kableExtra)
library(pander)
library(rmarkdown)
library(knitr)
library(stringr)
```

### Summary of the Functions

These are some functions I wrote primarily for use in creating tables. Rather than repeatedly tell it how to summarise the data and piece it together in a string, I let the function to do the main brunt of work. Being a manager instead of a worker has its perks, after all. 

In my opinion, these tend to work best when used in tandem with dplyr `summarise` function with grouped data. 

* `summ_interval` takes in continuous variables and outputs an estimate of centrality (your choice), along with an interval estimate (your choice), rounded to a number of digits (your choice), and seperated by a given delimeter (also your choice!). 

* `summ_prop` takes in either a logical vector, or a character vector with a supplied category to compare to and counts the numebr of matches along with a proportion. If a logical vector is supplied, no category is needed. This can be output as percentage or a percent, and with or without a $\%$ sign. Of note as well, if you want to treat $0 (0)$ as too uninterseting to matter, you can have it return a single '-' instead. Currently in dev is to handle factors more approporiately.

* `summ_na` works similarly to `summ_prop` except it works for counting tallying missingness. 

* `summ_full` is an extension of st_comb_IQR that also includes the range. It's not particularly useful, but I needed it for a table once.



```{r}
## Group Iris data by Species, and report summary statistics according to 
## different summ_interval input. Then output table with kableExtra

tab = iris %>% dplyr::group_by(Species) %>% 
  dplyr::summarise(Length1 = summ_interval(Sepal.Length, fun = mean),
                   Length2 = summ_interval(Sepal.Length, fun = median,
                                          limits = c(0.2, 0.6), digits = 0),
                   Length3 = summ_interval(Sepal.Length, fun = median,
                                          limits = 0:1, delim = "-"),
                   Length4 = summ_interval(Sepal.Length, fun = mean,
                                          limits = c(0.4, 0.6), delim = " ; ")) %>% t %>% 
  tibble::as.tibble()

##Format Table more cleanly
colnames(tab) = tab[1,]
tab = tab[-1,]

tab = tab %>% dplyr::mutate(summ_interval = 1:4) %>% 
  dplyr::select(c(4,1:3))

knitr::kable(tab,format = 'html') %>% 
  kableExtra::kable_styling(bootstrap_options = "hover")
```

```{r }
##Check Count and proportion of observations in mt cars with certain number of 
##cylinders, using logical vectors, and character vector with category

mtcars = mtcars %>% 
  dplyr::mutate(cyl_chr = as.character(cyl))

mtcars %>% dplyr::summarise(Cyl1 = summ_prop(cyl == 4, out = "percent"),
                            Cyl2 = summ_prop(cyl_chr, "4", out = "percent"),
                            Cyl3 = summ_prop(cyl_chr %in% c("4","8"),
                                                 out = "percent", perc_disp = T)) %>% 
  t %>% knitr::kable(format = 'html')  %>% 
  kableExtra::kable_styling(bootstrap_options = "hover")

```

```{r }
##Group mtcars by number of cylinders, and then summarise MPG, Count w prop 
## of Automatic Transmission, and Horsepower

mtcars %>% dplyr::group_by(cyl) %>% 
  dplyr::summarise(mpg = summ_interval(mpg),
                   am  = summ_prop(am == 1),
                   hp  = summ_interval(log2(hp+1), digits = 1,
                                      limits = 0:1)) %>% 
  t %>% knitr::kable(format = 'html')  %>% 
  kableExtra::kable_styling(bootstrap_options = "hover")
  
```


```{r}

X = 1:100
na.ind = sample(1:100, 15, replace = F)

summ_na(X)
summ_na(X, zero2dash = F)

X[na.ind] = NA

summ_na(X)
summ_na(X, out = "percentage")


```

