---
title: "freq_list, WIP v1"
author: "Andrew Frick"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{freq_list}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(ajfhelpR)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

freq_list is a function that replicates the functionality of SAS Proc Freq with List option set to true. There are currently some workarounds I'm using in the fucntion that end up being a teensy bit of a nuisance, but they should be remedied soon enough once I get around to it. Specifically, I dislike the `attach` workaround, and aim on fixing it.

Currently, this function accepts an n-dimensional table as input, and then tabulates every full combination of values observed, calculates the frequency and cumulative frequencies, while removing any combinations with frequency of 0.

Cumulative frequency is currently calculated with the rounded frequencies, so rounding errors are occuring. On the short list of fixes

```{r}
##Tabulate difference combinations amongst the mtcars dataset
library(stringr)
library(ajfhelpR)
library(tidyverse)

attach(mtcars)
freq_list(table(cyl,vs,am,gear,carb))

# Same as previous, but using head() to extract 5 most frequent combinations
head(freq_list(table(cyl,vs,am,gear,carb)))

freq_list(table(cyl,vs,am))
detach(mtcars)
```



```{r}
##tabulate categorical characteristics of diamonds dataset

attach(diamonds)

head(freq_list(table(cut,color,clarity)))

detach(diamonds)
```


